Adapting the MVDR to a GPU for active sonar systems of different sizes.

### 2013-01-22 ASA2013/asa2013.tex
la2pdf commit - 2013:01:22	22:22:37	(Tue22.Jan)

@@

If (\ref{spatialR}) were to be implemented directly, then we note that the estimation step would have a complexity of O($N_K N_L L^2$). By comparison performing a Gauss Jordan inversion is generally of O($L^3$), which implies that whenever $N_K\,N_L = N_K(M-L+1) \gg L$ then the estimation step will dominate. This tend to be true in active systems because it is usually required to set $L<M/2$ and $N_K\ge>3$ to MVDR yield the expected results.

@@

>>>>

To summarize as in Figure \ref{mvdr_beamforming}: For every pixel in the image we estimate $\eR$ (\ref{spatialR}), solve the linear equation $\eRi\1$ (\ref{weights}), the finally substitute some to get the beamformer output (\ref{mvdr_def}-\ref{weights}).

@@

If (\ref{spatialR}) were to be implemented directly, then we note that the estimation step would have a complexity of O($N_K N_L L^2$). By comparison, performing a Gauss Jordan inversion is generally of O($L^3$), which implies that whenever $N_K\,N_L = N_K(M-L+1) \gg L$ the estimation step will dominate. This tend to be true in active systems because we often want to set set $L<M/2$ and $N_K\ge>3$ for the MVDR to be robust enough to yield the expected results.

The good news is that this step can greatly benefit from both arithmetic optimizations and parallel computing power. By carefully evaluating (\ref{spatialR}) we can see that:
\begin{itemize}
\item 
\end{itemize}

@@

Some key aspects of our time domain MVDR implementation is illustrated in Figure \ref{mvdr_implementation}.

First we make each SM compute entire range lines of pixels, then load the subset of $\x$ that all these pixels depend upon into shared memory. This lets us perform temporal averaging without the need for reading additional channel data.

Second we assign $L$ threads per pixel to traverse the diagonals of $\eR$. On the top row a full computation is carried out, then that entire row is saved back to global memory following a collective access pattern that maximize global memory bandwidth. For subsequent rows the threads move along the diagonals while performing iterative summations; the result from the previous element on the diagonal is updated by adding and removing the correlation coefficients that enters and exits the sum, respectively. To minimize memory consumption, we compute the coefficients again every time we need them. When a thread has finished up a diagonal, we have them wrap around to compute one of the diagonals in the lower triangular of $\eR$. Since $\eR$ is conjugate symmetric, the values in the leftmost column is obtained by a complex conjugate copy of the relevant value in the first row. Combined these steps balances the load evenly on all threads, is almost completely free of arithmetic redundancy, and consumes less memory.

These challenges are very closely linked. The Quadro 6000 architecture is of compute capability 2.0, which means there are 48\,kB of shared memory (L1 cache) and 128\,kB of registers per SM~\cite{Nvidia2012}. This memory is shared by all active threads on that SM, a number that should be no less than 768~\cite{Nvidia2012a}. This will expose a sufficient level of data parallelism to ensure that memory latency is completely hidden (Little's law). By dividing the shared memory evenly on all 768 threads we find that each thread should store no more than 8 single precision complex numbers in shared memory\todo{introduce memory structure -shared for data shared between threads}, and 24 stored in registers. This should make it apparent why computing a single pixel per thread is a bad idea, and why we need to keep each thread as light on memory consumption as possible.
